---
title: "Comparing regressions, spatial regressions and machine learning techniques"
format: html
editor: source
---


see: 

**Spatial Models or Random Forest? Evaluating the Use of Spatially Explicit Machine Learning Methods to Predict Employment Density around New Transit Stations in Los Angeles** 

Kevin Credit 
*Geographical Analysis (2022) 54, 58–83* 



```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

packages <- c(
  "MASS",
  "tidyverse",
  "sf",
  "tmap",
  "here",
  "spdep", # areal data
  "grid",
  "GWmodel",
  "kableExtra",
  "cowplot",
  "parlitools",
  "corrplot",
  "lme4",
  "mgcv",
  "ggfortify",
  "ggforce", # has geom_circle
  "gridExtra",
  "car",
  "rgeoda", # for LISA
  "ggpubr",
  "ggrepel",
  "ggpointdensity",
  "ggalt", # for geom_encircle
  "ggsflabel", # has repel labels for sf
  "viridis",
  "spatialreg",
  "rgdal",
  "rgeos",
  "randomForest",
  "SpatialML" # spatial RF
  )


# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))

tmap_mode("plot")
rm(list=ls())
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

# avoid conflict with dplyr and MASS::select
select <- dplyr::select

set.seed(123)
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

df <- readRDS(here("df_withtext.rds")) |> 
  mutate(degree_educated = degree,
         professionals = industry_professional,
         younger_adults = age_18_to_19 + age_20_to_24 + age_25_to_29,
         english_speaking = english_all,
         single_ethnicity_household = 100 - ethnicity_mixed,
         not_good_health = health_fair + health_bad + health_very_bad,
         white = ethnicity_white,
         christian = christian,
         own_home = house_owned,
         no_car = cars_none,
         density = density)
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

# create neighbours list: polygon to nb
nb <- poly2nb(df |> st_buffer(dist=2000), queen = TRUE)
names(nb) <- factor(df$constituency_name)

# create weights: nb to listw
wts <- nb2listw(nb, zero.policy = TRUE)
names(wts$neighbours) <- factor(df$constituency_name)
names(wts$weights) <- factor(df$constituency_name)

```


```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

# moran_test <- moran.test(df$con_swing, wts)
# moran_test

```

## Ordinary Least Squares

### 1. OLS

```{r}

fmla <- con_swing ~ 
  degree_educated + 
  english_speaking + 
  not_good_health + 
  white + 
  LONG + 
  LAT

regOLS <-lm(fmla, df)
summary(regOLS)
```

Check residual spatial dependence

```{r}
lmMoranTest <- lm.morantest(regOLS, wts)
lmMoranTest

```



```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

# Model selection via LM tests

# lmLMtests <- lm.LMtests(regOLS, wts, test=c("LMerr", "LMlag", "RLMerr", "RLMlag", "SARMA"))
# lmLMtests
```

## Spatial Models 

### 2. SAR

```{r}

regSAR <- lagsarlm(fmla, data = df, wts)
summary(regSAR)

```

### 3. SEM

```{r}
regSEM <- errorsarlm(fmla, data=df, wts)
summary(regSEM)
```

### 4. SDM (Spatial Durbin Model)

add lag X to SAR

```{r}
regSDM <- lagsarlm(fmla, data = df, wts, type="mixed")
summary(regSDM)
```

### 5. SDEM (Spatial Durbin Error)

add lag X to SEM

```{r}
regSDEM <- errorsarlm(fmla, data = df, wts, etype = "emixed")
summary(regSDEM)
```

### 6. CAR

```{r}

regCAR1 <- gam(con_swing ~ degree_educated + 
                 english_speaking + 
                 not_good_health + 
                 white + 
                 LONG + 
                 LAT +
                 s(constituency_name,bs='mrf',xt=list(nb=nb),k=275),
               data=df, family=gaussian, weight=total_vote_19, method='REML')
summary(regCAR1)

```



```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 7. CAR2

# regCAR2 <- gam(con_swing ~ PC1 + PC2 + LONG + LAT +
#                 s(constituency_name,bs='mrf',xt=list(nb=nb),k=150, by = PC2) + 
#                 s(constituency_name,bs='mrf',xt=list(nb=nb),k=250),
#               data=df, family=gaussian, weight=total_vote_19, method='REML')
# summary(regCAR2)

```

```{r}
# results <- tibble(
#   indep_vars = c("degree_educated", "english_speaking", "not_good_health", "white","LONG","LAT","lagdegree_educated","lagenglish_speaking","lagnot_good_health",
#                  "lagwhite","lagLONG","lagLAT","lagY","rho","lambda","AIC"),
#   OLS = c(coef(regOLS)["degree_educated"],coef(regOLS)["english_speaking"],coef(regOLS)["not_good_health"],coef(regOLS)["white"],coef(regOLS)["LONG"],coef(regOLS)["LAT"],0,0,0,0,0,0,0,0,0,AIC(regOLS)),
#   SAR = c(coef(regSAR)["degree_educated"],coef(regSAR)["english_speaking"],coef(regSAR)["not_good_health"],coef(regSAR)["white"],coef(regSAR)["LONG"],coef(regSAR)["LAT"],0,0,0,0,0,0,coef(regSAR)["rho"],0,0,AIC(regSAR)),
#   SEM = c(coef(regSEM)["degree_educated"],coef(regSEM)["english_speaking"],coef(regSEM)["not_good_health"],coef(regSEM)["white"],coef(regSEM)["LONG"],coef(regSEM)["LAT"],0,0,0,0,0,0,0,coef(regSEM)["lambda"],0,AIC(regSEM)),
#   SDM = c(coef(regSDM)["degree_educated"],coef(regSDM)["english_speaking"],coef(regSDM)["not_good_health"],coef(regSDM)["white"],coef(regSDM)["LONG"],coef(regSDM)["LAT"],coef(regSDM)["lag.degree_educated"],coef(regSDM)["lag.english_speaking"],coef(regSDM)["lag.not_good_health"],coef(regSDM)["lag.white"],coef(regSDM)["lag.LONG"],coef(regSDM)["lag.LAT"],0,coef(regSDM)["rho"],0,AIC(regSDM)),
#   SDEM = c(coef(regSDEM)["degree_educated"],coef(regSDEM)["english_speaking"],coef(regSDEM)["not_good_health"],coef(regSDEM)["white"],coef(regSDEM)["LONG"],coef(regSDEM)["LAT"],coef(regSDEM)["lag.degree_educated"],coef(regSDEM)["lag.english_speaking"],coef(regSDEM)["lag.not_good_health"],coef(regSDEM)["lag.white"],coef(regSDEM)["lag.LONG"],coef(regSDEM)["lag.LAT"],0,0,coef(regSDEM)["lambda"],AIC(regSDEM)),
#   CAR1 = c(coef(regCAR1)["degree_educated"],coef(regCAR1)["english_speaking"],coef(regCAR1)["not_good_health"],coef(regCAR1)["white"],coef(regCAR1)["LONG"],coef(regCAR1)["LAT"],0,0,0,0,0,0,0,0,0,AIC(regCAR1))
# )
# results[,2:7] <- round(results[,2:7], 3)
```

## Random Forest Models

### 7. RF

```{r}
set.seed(123)
rf.fit <- randomForest(fmla, data=df, ntree=1500, 
                       keep.forest=FALSE, importance=TRUE)

#rf.fit

# Get variable importance from the model fit
ImpData.rf <- as.data.frame(randomForest::importance(rf.fit))

```


### 8. RFSAR

First add spatially lagged variables to `df`

```{r}
df$lag.degree_educated <- lag.listw(wts,df$degree_educated)
df$lag.english_speaking <- lag.listw(wts,df$english_speaking)
df$lag.not_good_health <- lag.listw(wts,df$not_good_health)
df$lag.white <- lag.listw(wts,df$white)
df$lag.LONG <- lag.listw(wts,df$LONG)
df$lag.LAT <- lag.listw(wts,df$LAT)
df$lag.con_swing <- lag.listw(wts,df$con_swing)
```

```{r}
rfsar.fit <- randomForest(con_swing ~ 
                            degree_educated + 
                            english_speaking + 
                            not_good_health + 
                            white + 
                            LONG + 
                            LAT +
                            lag.con_swing, data=df, ntree=1500, 
                       keep.forest=FALSE, importance=TRUE)

#rfsar.fit

# Get variable importance from the model fit
ImpData.rfsar <- as.data.frame(randomForest::importance(rfsar.fit))
```

### 9. RFSDM

```{r}

rfsdm.fit <- randomForest(con_swing ~ 
                            degree_educated + 
                            english_speaking + 
                            not_good_health + 
                            white + 
                            LONG + 
                            LAT + 
                            lag.degree_educated +
                            lag.english_speaking +
                            lag.not_good_health +
                            lag.white + 
                            lag.LONG +
                            lag.LAT +
                            lag.con_swing, 
                          data=df, ntree=1500, keep.forest=FALSE, importance=TRUE)

#rfsdm.fit

# Get variable importance from the model fit
ImpData.rfsdm <- as.data.frame(randomForest::importance(rfsdm.fit))

```

### 10. RFSLX

```{r}

rfslx.fit <- randomForest(con_swing ~ 
                            degree_educated + 
                            english_speaking + 
                            not_good_health + 
                            white + 
                            LONG + 
                            LAT + 
                            lag.degree_educated +
                            lag.english_speaking +
                            lag.not_good_health +
                            lag.white + 
                            lag.LONG +
                            lag.LAT, 
                          data=df, ntree=1500, keep.forest=FALSE, importance=TRUE)

#rfslx.fit

# Get variable importance from the model fit
ImpData.rfslx <- as.data.frame(randomForest::importance(rfslx.fit))

```


## Random Forest + rotation

### Find optimal rotation

Rotate by 1/10 of pi/2 each time and calculate MAE

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

df_ngrid <- df |> 
  st_transform(crs=27700) |> 
  mutate(centroid = st_centroid(geometry))

error_df <- data.frame(angle=seq(0,pi/2,pi/2/10),
                       error=rep(0,11))

coords <- st_coordinates(df_ngrid$centroid)

df_ngrid_nogeo <- df_ngrid |> st_drop_geometry()

for (i in 1:nrow(error_df)){
zzz <- df_ngrid_nogeo |> 
  mutate(geometry=df$geometry) |> 
  st_as_sf() |> 
  st_transform(crs = 27700) |> 
  select(con_swing,degree_educated, 
                 english_speaking, 
                 not_good_health, 
                 white) |> 
  st_drop_geometry()

zzz <- cbind(zzz,Rotation(coords,error_df[i,1]))
zzz <- zzz |> 
  rename(X="1",
         Y="2")
d.tree_loop <- randomForest(con_swing ~ .,
                data = zzz, ntree=1500, 
                       keep.forest=FALSE, importance=TRUE)

tree_est <- d.tree_loop$predicted
error_df[i,2] <- sum(abs(tree_est - df_ngrid$con_swing)) / nrow(df_ngrid)
}

error_df <- error_df |> 
  mutate(minMAE = case_when(error == min(error_df$error) ~ "min",
                            TRUE ~ "other"))
ggplot(error_df)+geom_point(aes(x=angle,y=error,colour=minMAE,size=minMAE)) + 
  # geom_smooth(aes(x=angle,y=error)) +
          scale_x_continuous(breaks  = c(seq(0, pi/2, pi/8)), 
                    labels = c("0", "\u03c0/8", "\u03c0/4", "3\u03c0/8", "\u03c0/2")) + 
  scale_colour_manual(values = c("red","black")) +
  scale_size_manual(values = c(2,0.25))+
  theme_bw() + 
  labs(title = paste0("MAE of regression tree: min at ",round(error_df[which.min(error_df$error),1],3)),
       subtitle = "for rotation from 0 to \u03c0")

```

### 11. RFr

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

coords <- st_coordinates(df_ngrid$centroid)

df_ngrid_rot <- cbind(df_ngrid,Rotation(coords,error_df[which.min(error_df$error),1]))

df_ngrid_rot <- df_ngrid_rot |> 
  mutate(LONG=X1,
         LAT=X2)

ggplot(df_ngrid_rot) + geom_point(aes(x=LONG, y=LAT)) + 
  geom_sf(lwd=0.1) + 
  coord_sf(datum = NA) + 
  theme_bw() + 
  labs(title = "Centroids",
       subtitle = "Optimally rotated for minimum absolute squared error")

fmla <- con_swing ~ 
  degree_educated + 
  english_speaking + 
  not_good_health + 
  white + 
  LONG + 
  LAT

set.seed(123)
rfr.fit <- randomForest(fmla, data=df_ngrid_rot, ntree=1500, 
                       keep.forest=FALSE, importance=TRUE)

#rf.fit

# Get variable importance from the model fit
ImpData.rfr <- as.data.frame(randomForest::importance(rfr.fit))

```


### 12. RFrSAR

First add spatially lagged variables to `df`

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

df_ngrid_rot$lag.degree_educated <- lag.listw(wts,df$degree_educated)
df_ngrid_rot$lag.english_speaking <- lag.listw(wts,df$english_speaking)
df_ngrid_rot$lag.not_good_health <- lag.listw(wts,df$not_good_health)
df_ngrid_rot$lag.white <- lag.listw(wts,df$white)
df_ngrid_rot$lag.LONG <- lag.listw(wts,df$LONG)
df_ngrid_rot$lag.LAT <- lag.listw(wts,df$LAT)
df_ngrid_rot$lag.con_swing <- lag.listw(wts,df$con_swing)
```

```{r}
rfrsar.fit <- randomForest(con_swing ~ 
                            degree_educated + 
                            english_speaking + 
                            not_good_health + 
                            white + 
                            LONG + 
                            LAT +
                            lag.con_swing, data=df_ngrid_rot, ntree=1500, 
                       keep.forest=FALSE, importance=TRUE)

#rfsar.fit

# Get variable importance from the model fit
ImpData.rfrsar <- as.data.frame(randomForest::importance(rfrsar.fit))
```

### 13. RFrSDM

```{r}

rfrsdm.fit <- randomForest(con_swing ~ 
                            degree_educated + 
                            english_speaking + 
                            not_good_health + 
                            white + 
                            LONG + 
                            LAT + 
                            lag.degree_educated +
                            lag.english_speaking +
                            lag.not_good_health +
                            lag.white + 
                            lag.LONG +
                            lag.LAT +
                            lag.con_swing, 
                          data=df_ngrid_rot, ntree=1500, keep.forest=FALSE, importance=TRUE)

#rfsdm.fit

# Get variable importance from the model fit
ImpData.rfrsdm <- as.data.frame(randomForest::importance(rfrsdm.fit))

```

### 14. RFrSLX

```{r}

rfrslx.fit <- randomForest(con_swing ~ 
                            degree_educated + 
                            english_speaking + 
                            not_good_health + 
                            white + 
                            LONG + 
                            LAT + 
                            lag.degree_educated +
                            lag.english_speaking +
                            lag.not_good_health +
                            lag.white + 
                            lag.LONG +
                            lag.LAT, 
                          data=df_ngrid_rot, ntree=1500, keep.forest=FALSE, importance=TRUE)

#rfslx.fit

# Get variable importance from the model fit
ImpData.rfrslx <- as.data.frame(randomForest::importance(rfrslx.fit))

```

## Mixed Models

### 15. LMER Mixed model

```{r}

mod_reduced2 <- lmer(con_swing ~ 
                       degree_educated + 
                       english_speaking + 
                       not_good_health + 
                       white + 
                       (1|region) +
                       (degree_educated|region) + 
                       #(english_speaking|region) + 
                       (not_good_health|region) +
                       (white|region),
                     df)

summary(mod_reduced2)
aic_mod_reduced2 <- AIC(mod_reduced2)

# ranef(mod_reduced2)

# these coefficients are, as they should be, the same as achieved in 'm3gam' further down
```


### 16. GAM mixed model with smoothing

```{r}
df |> st_intersects() |> 
  map_int(~length(.x)) %>% {
    which(. == 1)} -> temp
bufs <- rep(0,nrow(df))
bufs[temp] <- 2000 # just enough to give Isle of Wight neighbours
nlistconst <- df |> st_buffer(dist=bufs) |> 
  st_intersects() |> imap(~setdiff(.x,.y))
names(nlistconst) <- df$constituency_name

m3_gam<- gam(con_swing ~ 
               degree_educated + 
               english_speaking + 
               not_good_health + 
               white + 
               s(region, bs="re") + 
               s(degree_educated,region, bs="re") + 
               #s(english_speaking,region,  bs="re") + # not signif
               s(not_good_health,region,  bs="re") + 
               s(white,region,  bs="re") +
               s(constituency_name,bs='mrf',xt=list(nb=nlistconst),k=285),
             data=df, method="REML")

summary(m3_gam)

```


```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

results2 <- tibble(
  indep_vars = c("degree_educated", "english_speaking", "not_good_health", "white","LONG","LAT","lag_degree_educated","lag_english_speaking","lag_not_good_health",
                 "lag_white","lag_LONG","lag_LAT","lag_Y","rho","lambda","AIC"),
  OLS = c(coef(regOLS)["degree_educated"],coef(regOLS)["english_speaking"],coef(regOLS)["not_good_health"],coef(regOLS)["white"],coef(regOLS)["LONG"],coef(regOLS)["LAT"],0,0,0,0,0,0,0,0,0,AIC(regOLS)),
  SAR = c(coef(regSAR)["degree_educated"],coef(regSAR)["english_speaking"],coef(regSAR)["not_good_health"],coef(regSAR)["white"],coef(regSAR)["LONG"],coef(regSAR)["LAT"],0,0,0,0,0,0,0,coef(regSAR)["rho"],0,AIC(regSAR)),
  SEM = c(coef(regSEM)["degree_educated"],coef(regSEM)["english_speaking"],coef(regSEM)["not_good_health"],coef(regSEM)["white"],coef(regSEM)["LONG"],coef(regSEM)["LAT"],0,0,0,0,0,0,0,0,coef(regSEM)["lambda"],AIC(regSEM)),
  SDM = c(coef(regSDM)["degree_educated"],coef(regSDM)["english_speaking"],coef(regSDM)["not_good_health"],coef(regSDM)["white"],coef(regSDM)["LONG"],coef(regSDM)["LAT"],coef(regSDM)["lag.degree_educated"],coef(regSDM)["lag.english_speaking"],coef(regSDM)["lag.not_good_health"],coef(regSDM)["lag.white"],coef(regSDM)["lag.LONG"],coef(regSDM)["lag.LAT"],0,coef(regSDM)["rho"],0,AIC(regSDM)),
  SDEM = c(coef(regSDEM)["degree_educated"],coef(regSDEM)["english_speaking"],coef(regSDEM)["not_good_health"],coef(regSDEM)["white"],coef(regSDEM)["LONG"],coef(regSDEM)["LAT"],coef(regSDEM)["lag.degree_educated"],coef(regSDEM)["lag.english_speaking"],coef(regSDEM)["lag.not_good_health"],coef(regSDEM)["lag.white"],coef(regSDEM)["lag.LONG"],coef(regSDEM)["lag.LAT"],0,0,coef(regSDEM)["lambda"],AIC(regSDEM)),
  CAR1 = c(coef(regCAR1)["degree_educated"],coef(regCAR1)["english_speaking"],coef(regCAR1)["not_good_health"],coef(regCAR1)["white"],coef(regCAR1)["LONG"],coef(regCAR1)["LAT"],0,0,0,0,0,0,0,0,0,AIC(regCAR1)),
  LMER = c(fixed.effects(mod_reduced2)[2],fixed.effects(mod_reduced2)[3],fixed.effects(mod_reduced2)[4],fixed.effects(mod_reduced2)[5],0,0,0,0,0,0,0,0,0,0,0,AIC(mod_reduced2)),
  GAMmixed = c(m3_gam$coefficients[2],m3_gam$coefficients[3],m3_gam$coefficients[4],m3_gam$coefficients[5],0,0,0,0,0,0,0,0,0,0,0,AIC(m3_gam)),
  RF = c(ImpData.rf[1,1],ImpData.rf[2,1],ImpData.rf[3,1],ImpData.rf[4,1],ImpData.rf[5,1],ImpData.rf[6,1],0,0,0,0,0,0,0,0,0,0),
  RFSAR = c(ImpData.rfsar[1,1],ImpData.rfsar[2,1],ImpData.rfsar[3,1],ImpData.rfsar[4,1],ImpData.rfsar[5,1],ImpData.rfsar[6,1],0,0,0,0,0,0,ImpData.rfsar[7,1],0,0,0),
  RFSDM = c(ImpData.rfsdm[1,1],ImpData.rfsdm[2,1],ImpData.rfsdm[3,1],ImpData.rfsdm[4,1],ImpData.rfsdm[5,1],ImpData.rfsdm[6,1],ImpData.rfsdm[7,1],ImpData.rfsdm[8,1],ImpData.rfsdm[9,1],ImpData.rfsdm[10,1],ImpData.rfsdm[11,1],ImpData.rfsdm[12,1],ImpData.rfsdm[13,1],0,0,0),
  RFSLX =c(ImpData.rfslx[1,1],ImpData.rfslx[2,1],ImpData.rfslx[3,1],ImpData.rfslx[4,1],ImpData.rfslx[5,1],ImpData.rfslx[6,1],ImpData.rfslx[7,1],ImpData.rfslx[8,1],ImpData.rfslx[9,1],ImpData.rfslx[10,1],ImpData.rfslx[11,1],ImpData.rfslx[12,1],0,0,0,0),
  RFr = c(ImpData.rfr[1,1],ImpData.rfr[2,1],ImpData.rfr[3,1],ImpData.rfr[4,1],ImpData.rfr[5,1],ImpData.rfr[6,1],0,0,0,0,0,0,0,0,0,0),
  RFrSAR = c(ImpData.rfrsar[1,1],ImpData.rfrsar[2,1],ImpData.rfrsar[3,1],ImpData.rfrsar[4,1],ImpData.rfrsar[5,1],ImpData.rfrsar[6,1],0,0,0,0,0,0,ImpData.rfrsar[7,1],0,0,0),
  RFrSDM = c(ImpData.rfrsdm[1,1],ImpData.rfrsdm[2,1],ImpData.rfrsdm[3,1],ImpData.rfrsdm[4,1],ImpData.rfrsdm[5,1],ImpData.rfrsdm[6,1],ImpData.rfrsdm[7,1],ImpData.rfrsdm[8,1],ImpData.rfrsdm[9,1],ImpData.rfrsdm[10,1],ImpData.rfrsdm[11,1],ImpData.rfrsdm[12,1],ImpData.rfrsdm[13,1],0,0,0),
  RFrSLX =c(ImpData.rfrslx[1,1],ImpData.rfrslx[2,1],ImpData.rfrslx[3,1],ImpData.rfrslx[4,1],ImpData.rfrslx[5,1],ImpData.rfrslx[6,1],ImpData.rfrslx[7,1],ImpData.rfrslx[8,1],ImpData.rfrslx[9,1],ImpData.rfrslx[10,1],ImpData.rfrslx[11,1],ImpData.rfrslx[12,1],0,0,0,0)
)

errors <- tibble(
  metric = c("R^2",
             #"Adj R^2",
             "MSE",
             "RMSE",
             "varSE"),
  OLS = rep(0,4),
  SAR = rep(0,4),
  SEM = rep(0,4),
  SDM = rep(0,4),
  SDEM = rep(0,4),
  CAR1 = rep(0,4),
  #CAR2 = rep(0,4),
  LMER = rep(0,4),
  GAMmixed= rep(0,4),
  RF = rep(0,4),
  RFSAR = rep(0,4),
  RFSDM = rep(0,4),
  RFSLX = rep(0,4),
  RFr = rep(0,4),
  RFrSAR = rep(0,4),
  RFrSDM = rep(0,4),
  RFrSLX = rep(0,4)
)

```


```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 1. OLS

errors$OLS <- c(1 - (sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/564)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - regOLS$fitted.values)^2),
                    sqrt(mean((df$con_swing - regOLS$fitted.values)^2)),
                    var((df$con_swing - regOLS$fitted.values)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 2. SAR

errors$SAR <- c(1 - (sum((df$con_swing-regSAR$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/563)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - regSAR$fitted.values)^2),
                    sqrt(mean((df$con_swing - regSAR$fitted.values)^2)),
                    var((df$con_swing - regSAR$fitted.values)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 3. SEM

errors$SEM <- c(1 - (sum((df$con_swing-regSEM$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/564)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - regSEM$fitted.values)^2),
                    sqrt(mean((df$con_swing - regSEM$fitted.values)^2)),
                    var((df$con_swing - regSEM$fitted.values)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 4. SDM

errors$SDM <- c(1 - (sum((df$con_swing-regSDM$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/558)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - regSDM$fitted.values)^2),
                    sqrt(mean((df$con_swing - regSDM$fitted.values)^2)),
                    var((df$con_swing - regSDM$fitted.values)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 5. SDEM

errors$SDEM <- c(1 - (sum((df$con_swing-regSDEM$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/558)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - regSDEM$fitted.values)^2),
                    sqrt(mean((df$con_swing - regSDEM$fitted.values)^2)),
                    var((df$con_swing - regSDEM$fitted.values)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 6. CAR1

errors$CAR1 <- c(1 - (sum((df$con_swing-regCAR1$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/564)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - regCAR1$fitted.values)^2),
                    sqrt(mean((df$con_swing - regCAR1$fitted.values)^2)),
                    var((df$con_swing - regCAR1$fitted.values)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 7. CAR2

# errors$CAR2 <- c(1 - (sum((df$con_swing-regCAR2$predicted)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
#                  mean((df$con_swing - regCAR2$fitted.values)^2),
#                  sqrt(mean((df$con_swing - regCAR2$fitted.values)^2)),
#                  var((df$con_swing - regCAR2$fitted.values)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 8. RF

errors$RF <- c(1 - (sum((df$con_swing-rf.fit$predicted)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/564)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - rf.fit$predicted)^2),
                    sqrt(mean((df$con_swing - rf.fit$predicted)^2)),
                    var((df$con_swing - rf.fit$predicted)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 9. RFSAR

errors$RFSAR <- c(1 - (sum((df$con_swing-rfsar.fit$predicted)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/563)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - rfsar.fit$predicted)^2),
                    sqrt(mean((df$con_swing - rfsar.fit$predicted)^2)),
                    var((df$con_swing - rfsar.fit$predicted)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 10. RFSDM

errors$RFSDM <- c(1 - (sum((df$con_swing-rfsdm.fit$predicted)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/557)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                  mean((df$con_swing - rfsdm.fit$predicted)^2),
                    sqrt(mean((df$con_swing - rfsdm.fit$predicted)^2)),
                    var((df$con_swing - rfsdm.fit$predicted)^2))

```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 11. RFSLX

errors$RFSLX <- c(1 - (sum((df$con_swing-rfslx.fit$predicted)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/558)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                  mean((df$con_swing - rfslx.fit$predicted)^2),
                    sqrt(mean((df$con_swing - rfslx.fit$predicted)^2)),
                    var((df$con_swing - rfslx.fit$predicted)^2))
```


```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 12. RFr

errors$RFr <- c(1 - (sum((df$con_swing-rfr.fit$predicted)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/564)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - rfr.fit$predicted)^2),
                    sqrt(mean((df$con_swing - rfr.fit$predicted)^2)),
                    var((df$con_swing - rfr.fit$predicted)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 13. RFrSAR

errors$RFrSAR <- c(1 - (sum((df$con_swing-rfrsar.fit$predicted)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/563)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                mean((df$con_swing - rfrsar.fit$predicted)^2),
                    sqrt(mean((df$con_swing - rfrsar.fit$predicted)^2)),
                    var((df$con_swing - rfrsar.fit$predicted)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 14. RFrSDM

errors$RFrSDM <- c(1 - (sum((df$con_swing-rfrsdm.fit$predicted)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/557)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                  mean((df$con_swing - rfrsdm.fit$predicted)^2),
                    sqrt(mean((df$con_swing - rfrsdm.fit$predicted)^2)),
                    var((df$con_swing - rfrsdm.fit$predicted)^2))

```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 15. RFrSLX

errors$RFrSLX <- c(1 - (sum((df$con_swing-rfrslx.fit$predicted)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/558)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                  mean((df$con_swing - rfrslx.fit$predicted)^2),
                    sqrt(mean((df$con_swing - rfrslx.fit$predicted)^2)),
                    var((df$con_swing - rfrslx.fit$predicted)^2))
```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 16. RFrSDM

errors$LMER <- c(1 - (sum((df$con_swing-fitted(mod_reduced2))^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/557)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                  mean((df$con_swing - fitted(mod_reduced2))^2),
                    sqrt(mean((df$con_swing - fitted(mod_reduced2))^2)),
                    var((df$con_swing - fitted(mod_reduced2))^2))

```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

### 17. RFrSLX

errors$GAMmixed <- c(1 - (sum((df$con_swing-m3_gam$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                #1 - (570/558)*(sum((df$con_swing-regOLS$fitted.values)^2)/sum((df$con_swing-mean(df$con_swing))^2)),
                  mean((df$con_swing - m3_gam$fitted.values)^2),
                    sqrt(mean((df$con_swing - m3_gam$fitted.values)^2)),
                    var((df$con_swing - m3_gam$fitted.values)^2))
```

## All results mixed / non-mixed models

```{r, fig.width=12}
#| code-fold: true

err_res <- rbind(
  results2 |> rename(vars_metric = indep_vars),
  errors |> rename(vars_metric = metric)
  ) |> 
  mutate(across(where(is.numeric), round, digits=3)) |> 
  na_if(0) 

options(knitr.kable.NA = "")  



err_res |>
  kbl() |> 
  add_header_above(c(" "=2, "Coefficients"=7,"Permutation importances"=8)) |> 
  add_header_above(c(" "=2, "'spatialreg' Regressions"=4,"CAR"=1,"Mixed Models"=2,"Random Forests"=4,"Rotated RF"=4)) |> 
  pack_rows("Indep_Vars (coefficient / variable importance)", 1, 6) |> 
  pack_rows("Lagged_Vars (coefficient / variable importance)", 7, 13) |> 
  pack_rows("Spatial_Params", 14, 15) |> 
  pack_rows("Accuracy", 16, 20) |> 
  column_spec(10:13, color = "darkgreen") |> 
  column_spec(14:17, color = "blue") |> 
  column_spec(7, color = "red") |> 
  column_spec(8:9, color = "purple") |> 
  row_spec(17, bold=TRUE) |> 
  kable_minimal(full_width = T)

```


### Mapped residuals

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

df$residols <- regOLS$residuals
r1 <- ggplot(df)+geom_sf(aes(fill=residols), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals OLS")

df$residsar <- regSAR$residuals
r2 <- ggplot(df)+geom_sf(aes(fill=residsar), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals SAR")

df$residsem <- regSEM$residuals
r3 <- ggplot(df)+geom_sf(aes(fill=residsem), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals SEM")

df$residsdm <- regSDM$residuals
r4 <- ggplot(df)+geom_sf(aes(fill=residsdm), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals SDM")

df$residsdem <- regSDEM$residuals
r5 <- ggplot(df)+geom_sf(aes(fill=residsdem), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals SDEM")

df$residcar1 <- regCAR1$residuals
r6 <- ggplot(df)+geom_sf(aes(fill=residcar1), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals CAR1")

df$residrf <- df$con_swing - rf.fit$predicted
r8 <- ggplot(df)+geom_sf(aes(fill=residrf), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals RF")

df$residrfsar <- df$con_swing - rfsar.fit$predicted
r9 <- ggplot(df)+geom_sf(aes(fill=residrfsar), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals RFSAR")

df$residrfsdm <- df$con_swing - rfsdm.fit$predicted
r10 <- ggplot(df)+geom_sf(aes(fill=residrfsdm), lwd=0.05) +
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) +
  theme_bw() +
  labs(title = "Residuals RFSDM")

df$residrfslx <- df$con_swing - rfslx.fit$predicted
r11 <- ggplot(df)+geom_sf(aes(fill=residrfslx), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals RFSLX")

df$residrfr <- df$con_swing - rfr.fit$predicted
r12 <- ggplot(df)+geom_sf(aes(fill=residrfr), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals RFr")

df$residrfrsar <- df$con_swing - rfrsar.fit$predicted
r13 <- ggplot(df)+geom_sf(aes(fill=residrfrsar), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals RFrSAR")

df$residrfrsdm <- df$con_swing - rfrsdm.fit$predicted
r14 <- ggplot(df)+geom_sf(aes(fill=residrfrsdm), lwd=0.05) +
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) +
  theme_bw() +
  labs(title = "Residuals RFrSDM")

df$residrfrslx <- df$con_swing - rfrslx.fit$predicted
r15 <- ggplot(df)+geom_sf(aes(fill=residrfrslx), lwd=0.05) + 
  scale_fill_gradient2(low="red",high="blue",mid="white",midpoint=0) + 
  theme_bw() + 
  labs(title = "Residuals RFrSLX")
```


```{r}

resids_together <- ggarrange(r1, r2, r3, r4, r5, 
                             r6, NULL, NULL, NULL, NULL, 
                             r8, r9, r10, r11, NULL, 
                             r12, r13, r14, r15, NULL,
                             nrow=4, ncol=5, common.legend = TRUE, legend="bottom")

```

```{r, fig.width=12, fig.height=12}

resids_together

```








## Geographic RF

```{r}

Coords <- df |> 
  select(LONG,LAT) |> 
  st_drop_geometry() |> 
  as.matrix()

df.grf <- df |> 
  select(con_swing, degree_educated, english_speaking, not_good_health, white) |> 
  st_drop_geometry()

bwe <-grf.bw(con_swing ~ 
               degree_educated + 
               english_speaking + 
               not_good_health + 
               white,
             df.grf, kernel="adaptive",
             coords=Coords, bw.min = 20, bw.max = 100, step = 10,
             forests = FALSE, weighted = TRUE)

grf <- grf(con_swing ~ 
               degree_educated + 
               english_speaking + 
               not_good_health + 
               white, 
           dframe=df.grf, bw=bwe$Best.BW,
           kernel="adaptive", coords=Coords,
           forests = TRUE, weighted = TRUE)

```

```{r}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false


df$rflocalimp_degree_educated <- grf$Local.Variable.Importance$degree_educated
df$rflocalimp_english_speaking <- grf$Local.Variable.Importance$english_speaking
df$rflocalimp_not_good_health <- grf$Local.Variable.Importance$not_good_health
df$rflocalimp_white <- grf$Local.Variable.Importance$white

grfp1 <- ggplot(df) + 
  geom_sf(aes(fill=rflocalimp_degree_educated), lwd=0.05) + 
  scale_fill_viridis(limits=c(0,370), direction = -1,option = "E") + 
  labs(title = "degree_educated",
       subtitle = "Variable importance",
       fill="") + 
  theme_bw()

grfp2 <- ggplot(df) + 
  geom_sf(aes(fill=rflocalimp_english_speaking), lwd=0.05) + 
  scale_fill_viridis(limits=c(0,370), direction = -1,option = "E") +
  labs(title = "english_speaking",
       subtitle = "Variable importance",
       fill="") + 
  theme_bw()

grfp3 <- ggplot(df) + 
  geom_sf(aes(fill=rflocalimp_not_good_health), lwd=0.05) + 
  scale_fill_viridis(limits=c(0,370), direction = -1,option = "E") +
  labs(title = "not_good_health",
       subtitle = "Variable importance",
       fill="") + 
  theme_bw()

grfp4 <- ggplot(df) + 
  geom_sf(aes(fill=rflocalimp_white), lwd=0.05) + 
  scale_fill_viridis(limits=c(0,370), direction = -1,option = "E") +
  labs(title = "white",
       subtitle = "Variable importance",
       fill="") + 
  theme_bw()

grf_varimp_plots <- ggarrange(grfp1, grfp2, grfp3, grfp4, 
          nrow = 1, ncol = 4, legend = "bottom", common.legend = TRUE)

###

grfp1lon <- ggplot(df |> filter(region == "London")) +
  geom_sf(aes(fill=rflocalimp_degree_educated), lwd=0.05) + 
  scale_fill_viridis(limits=c(0,370), direction = -1,option = "E") + 
  labs(title = "degree_educated",
       subtitle = "Variable importance",
       fill="") + 
  theme_bw()

grfp2lon <- ggplot(df |> filter(region == "London")) +
  geom_sf(aes(fill=rflocalimp_english_speaking), lwd=0.05) + 
  scale_fill_viridis(limits=c(0,370), direction = -1,option = "E") +
  labs(title = "english_speaking",
       subtitle = "Variable importance",
       fill="") + 
  theme_bw()

grfp3lon <- ggplot(df |> filter(region == "London")) +
  geom_sf(aes(fill=rflocalimp_not_good_health), lwd=0.05) + 
  scale_fill_viridis(limits=c(0,370), direction = -1,option = "E") +
  labs(title = "not_good_health",
       subtitle = "Variable importance",
       fill="") + 
  theme_bw()

grfp4lon <- ggplot(df |> filter(region == "London")) +
  geom_sf(aes(fill=rflocalimp_white), lwd=0.05) + 
  scale_fill_viridis(limits=c(0,370), direction = -1,option = "E") +
  labs(title = "white",
       subtitle = "Variable importance",
       fill="") + 
  theme_bw()

grf_varimp_plotslon <- ggarrange(grfp1lon, grfp2lon, grfp3lon, grfp4lon, 
          nrow = 1, ncol = 4, legend = "bottom", common.legend = TRUE)
```

```{r, fig.width=10, fig.height=4}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false


grf_varimp_plots

```


```{r, fig.width=10, fig.height=4}
#| code-fold: false
#| message: false
#| warning: false
#| echo: false

grf_varimp_plotslon

```

```{r}
testdf <- df.grf |> 
  mutate(X=df$LONG,
         Y=df$LAT)

grf$Global.Model$prediction.error
```

